Fixing Vulnerabilities Potentially Hinders Maintainability

### 2021-02-09 response/response.tex

\begin{revcomment}[1.4]

    The authors could have adopted both approaches, Better 
    Code Hub guidelines or another one first and then measure 
    the code maintainability with the Delta Maintainability Model 
    since the goal of the paper is to improve security on the 
    maintainability.

\end{revcomment}

\Us As we explain above, namely while addressing comment 1.2, both models assess 
maintainability. We consider the Delta Maintainability 
Model to be an alternative to our model. However, it only uses 5 guidelines of the 
set of guidelines our study considers. This has been 
discussed in the related work section.

@@

\Us Thank you for bringing our attention to this paper. We now discuss
this paper in the Related Work section and used it to support some 
of our claims and findings.

We added the following paragraph to Section 7 (\emph{Related Work}):

\begin{quote}
    Researchers
performed a large-scale empirical study to understand the characteristics of security patches
and their differences against bug fixes [Li2017].
The main findings were that security patches are smaller and less complex compared
to bug fixes and usually performed at function-level. Our study compares
the impact of security patches on software maintainability with the impact 
of regular changes.
\end{quote}

We also used the paper to suport our research in different Sections of our paper:

Section 1 (\emph{Introduction})
\begin{quote}
    Our hypothesis is that some of these patches may 
have a negative impact on the software maintainability and, 
possibly, even be the cause of the introduction of new 
vulnerabilities---harming software reliability and introducing 
technical debt. Research found that $34\%$ of the security patches 
performed introduce new problems and $52\%$ are incomplete and do not 
fully secure systems [Li2017].
\end{quote}

Section 3.2 (\emph{Security Patches vs. Regular Changes})
\begin{quote}
    Previous studies attempted to measure the impact of regular changes 
on open-source software maintainability. 
However, there is no previous work focused on comparing the impact 
of security patches with regular changes on maintainability, only
with bug-fixes [Li2017].
\end{quote}

Section 5 (\emph{Study Implications})
\begin{quote}
    \textit{\textbf{Prioritize High and Medium Severity:}} Previous research 
exhibits proof that developers prioritize
higher impact vulnerabilities [Li2017].
Our study shows that vulnerabilities of high and
medium severity should be prioritized in software maintainability tasks.
\end{quote}

@@

\Us Thank you for pointing out all of these issues. We addressed all of them. We decided
to use regular changes/regular commits instead of baseline and non-security commits.

\begin{revcomment}[2.6]
    Section 3\\
    \begin{itemize}
        \item Sometime you say 1330 patches (e.g. page 7, line 16) and sometimes 1300 (e.g. page 8, line 11, 
        and the abstract).  In science it's better to say the exact number and use it always and not round.  
        \item You say the dataset had 1330 (or 1300) patches and 1282 commits though you also say one patch 
        can have multiple commits assigned.  So how did you end up with less than one commit/patch?
    \end{itemize}
\end{revcomment}

\Us Thank you for raising this issue. The right number of security patches is $1300$ ($624$ from Ponta et al. and 
$676$ from Secbench). The $1330$ value is a typo, we fixed the issue. Regarding the second point, $1282$ commits is
the equivalent to the $624$ patches that integrate the Pontas et al. dataset where one patch
can have multiple commits. The $1282$ commits are referring to the Ponta et al. dataset and 
not to the combined dataset.

\begin{quote}
    We use a combined dataset of $1300$ security patches which is 
the outcome of mining and manually inspecting a total of $312$ 
GitHub projects.
\end{quote}

### 2021-02-09 response/response.tex

\begin{quote}
BCH evaluates the codebase available in the default branch of a GitHub project. 
  We created a tool that pulls the codebase of each commit of our dataset 
  to a new branch; it sets the new branch as the default branch; and, runs 
  the BCH analysis on the codebase; after the analysis is finished, the tool saves 
  the BCH metrics results to a cache file.
\end{quote}

\begin{revcomment}[2.9]
    Section 3\\
    \begin{itemize}
        \item Please explain your repeatable methodology to "clean the dataset and select the most relevant and 
        compatible projects" (page 8, line 14) that brings the patches from 1282 to 969.  
        \item From 969 to 866 patches - 
        you just could not classify?  Please explain.
    \end{itemize}
\end{revcomment}

\Us As explained in comment $2.6$, the $1282$ commits refer only to the Ponta et al. dataset (a total of $624$ 
patches). In total, we analyzed $1300$ patches ($624$ patches from Ponta et al. and $676$ from Secbench). 
However, our study only considers $969$ patches. We detected a total of $23$ patches with floss-refactorings. 
In addition, BCH has some limitations, in particular, the lack of language support and project size,
and was uncapable of analyzing the rest of the $308$ data points. All of this is explained
in more detail later in Section 3.4.

\begin{quote}
    Security patches can be performed through one commit (single-commit); 
    several consecutive commits (multi-commits); or, commit(s) interleaved with more 
    programming activities (floss-refactoring). Only $10.7\%$ of the data points 
    of our dataset involve more than one commit, the other
    $89.3\%$ of the cases are single-commit patches. We manually 
    inspected $25.1\%$ ($244$/$969$) of the security patches---$122$ from each dataset---and, 
    identified a total of $23$ floss-refactorings. Many of these 
    are patches with many changes where it is hard to understand which parts 
    involve the security patch. These were disregarded to minimize the impact 
    of measuring other programming activities rather than solely security patches
    can have in our results.

    For projects with large codebases, the results retrieved for 
    \emph{Keep Your Codebase Small} were not viable because they 
    retrieved values above the limit set by BCH ($20$ Person-years). Thus, 
    we decided to not consider this guideline in our research. The 
    \emph{Automated Tests} guideline was also not considered since the 
    tool does not contemplate two of the most important techniques to 
    security testing: vulnerability scanning and penetration testing. 
    Instead, it only contemplates unit testing. Due to BCH limitations, 
    we detected $308$ data points that retrieved incorrect overall 
    calculations. Those data points were disregarded from the study.
\end{quote}

@@

\begin{revcomment}[2.13]
    Section 3\\
    \begin{itemize}
        \item Page 10, line 49 "mainly single-commit patches" … "small percentage 
        of data points" - give us the exact numbers for each of these.  
        \item Page 11, line 38 - define "floss refactoring" and justify how these were objectively 
        and repeatably identified in your set. You say "these cases" which makes 
        it seems that you were including the set of patches that had more than one 
        commit as your (only) floss candidates.  
    \end{itemize}
\end{revcomment}

\Us Our dataset integrates $89.3\%$ of single-commit patches and 
$10.7\%$ of patches involving more than one commit. We replaced 
line 49 by the following sentence:

\begin{quote}
    Only $10.7\%$ of the data points 
of our dataset involve more than one commits, the other
$89.3\%$ of the cases are single-commit patches.
\end{quote}

We define floss-refactoring with the paragraph below. 
Regarding the methodology to detect floss-refactorings, 
we manually inspected $25.1\%$ ($244$/$969$) of the security 
patches---$122$ from each dataset---and, 
identified a total of $23$ floss-refactorings.

\begin{quote}
    Security patches can be performed through one commit (single-commit); 
    several consecutive commits (multi-commits); or, commit(s) interleaved with more 
    programming activities (floss-refactoring). Only $10.7\%$ of the data points 
    of our dataset involve more than one commit, the other
    $89.3\%$ of the cases are single-commit patches. We manually 
    inspected $25.1\%$ ($244$/$969$) of the security patches---$122$ from each dataset---and, 
    identified a total of $23$ floss-refactorings. Many of these 
    are patches with many changes where it is hard to understand which parts 
    involve the security patch. These were disregarded to minimize the impact 
    of measuring other programming activities rather than solely security patches
    can have in our results. 
\end{quote}

@@

\begin{revcomment}[2.21]
    Section 4\\
    \begin{itemize}
        \item RQ3 is answered to succinctly and with too little 
        detail to be convincing to me that the differences between 
        security and regular changes is different.  
    \end{itemize}
\end{revcomment}

\Us We improved the discussion of the comparison between 
the results for security patches and regular changes. We now
consider two baselines: \emph{size-baseline}, where 
we collected random changes with the same size as security 
patches; and, \emph{random-baseline}, where we collected
random changes without any restrictions. The overall results
show that regular changes are less prone to hinder software 
maintainability than security patches. Security
patches results have more cases with negative 
impact on software maintainability than any baseline 
of regular changes. By providing results 
for both baselines, we now clearly see that the limitation 
found for the \emph{size-baseline} does not change 
the need to pay more attention to security patches 
in software maintenance tasks.

\begin{revcomment}[2.22]
    Section 5\\
    \begin{itemize}
        \item Line 32 - since RQ3 was not convincing, 
        I don't feel a separate effort for maintainable security 
        is justified as separate from general focus on maintainability. 
    \end{itemize}
\end{revcomment}

\Us As explained in the previous response, security patches 
hinder more software maintainability than regular changes with size 
restriction (\emph{size-baseline}) or no size restriction (\emph{random-baseline}). Thus, we argue that
special attention should be given to maintainable security
when teaching software maintainability. We improved the analysis 
provided in Section 4 for the research question 3 (RQ3). We added 
an extra baseline for comparison with no size restrictions (\emph{random-baseline}) and 
improved the discussion over our results.

### 2021-02-09 paper/emse20.tex
improving 2.13

The same approach is used to analyze the impact of regular 
changes. Security patches can be performed through one commit (single-commit); 
several consecutive commits (multi-commits); or, commit(s) interleaved with more 
programming activities (floss-refactoring). Only $10.7\%$ of the data points 
of our dataset involve more than one commit, the other
$89.3\%$ of the cases are single-commit patches. We manually 
inspected all the cases and identified $23$ floss-refactorings. 
These were disregarded since it would not be fair to measure the 
maintainability of these cases where other programming activities 
are involved.

@@

\textit{\textbf{Internal Validity:}} The security patches dataset provided by previous
work~\cite{Reis:2017:IJSSE} was collected based on the messages of GitHub
commits produced by project developers to classify the changes performed while 
patching vulnerabilities. This approach discards patches that were
not explicit in commits messages. We assume that patches were performed 
using a single-commit or several sequentially. The 
perspective that a 
developer may quickly perform a patch and later 
proceed to the refactor is not considered. We assume 
that all patches were only performed once. Depending 
on the impact of the vulnerability in the system, some 
vulnerabilities may have more urgency to be patched than 
others. For instance, a vulnerability performing a Denial-of-Service attack
which usually brings entire systems down may be more urgent to patch than a 
cross-site scripting vulnerability which usually does not have impact
on the systems execution but rather on the data accessibility.

@@

\textit{\textbf{External Validity:}} The BCH tool uses private and open-source data to 
determine the thresholds for each guideline. We only analyze patches of open-source software.
Thus, our findings may not extend to private/non-open source software. Different programming 
languages may require different coding practices to address software safety. The 
dataset comprises more commits in Java, i.e., the dataset may not be representative 
of the population regarding programming languages. For both datasets, we a manual validation 
of the message of the commits was performed. Only commits in English were considered. Thus, our approach does not consider 
patches in any other language but English.

>>>>

Security patches can be performed through one commit (single-commit); 
several consecutive commits (multi-commits); or, commit(s) interleaved with more 
programming activities (floss-refactoring). Only $10.7\%$ of the data points 
of our dataset involve more than one commit, the other
$89.3\%$ of the cases are single-commit patches. We manually 
inspected $25.1\%$ ($244$/$969$) of the security patches---$122$ from each dataset---and, 
identified a total of $23$ floss-refactorings. Many of these 
are patches with many changes where it is hard to understand which parts 
involve the security patch. These were disregarded to minimize the impact 
of measuring other programming activities rather than solely security patches
can have in our results.

@@

\textit{\textbf{Internal Validity:}} The security patches dataset provided by previous
work~\cite{Reis:2017:IJSSE} was collected based on the messages of GitHub
commits produced by project developers to classify the changes performed while 
patching vulnerabilities. This approach discards patches that were
not explicit in commits messages. We assume that patches were performed 
using a single-commit or several sequentially. The 
perspective that a 
developer may quickly perform a patch and later 
proceed to the refactor is not considered. We assume 
that all patches were only performed once. Depending 
on the impact of the vulnerability in the system, some 
vulnerabilities may have more urgency to be patched than 
others. For instance, a vulnerability performing a Denial-of-Service attack
which usually brings entire systems down may be more urgent to patch than a 
cross-site scripting vulnerability which usually does not have impact
on the systems execution but rather on the data accessibility.
We manually inspected $25.1\%$ of the security patches looking 
for floss-refactorings---$122$ from each dataset. We did 
find $23$ cases we argue to be floss-refactorings and toss them
in order to minimize the impact of this threat.

@@

\textit{\textbf{External Validity:}} The BCH tool uses private and open-source data to 
determine the thresholds for each guideline. We only analyze patches of open-source software.
Thus, our findings may not extend to private/non-open source software. Different programming 
languages may require different coding practices to address software safety. The 
dataset comprises more commits in Java, i.e., the dataset may not be representative 
of the population regarding programming languages. For both datasets, a manual validation 
of the message of the commits was performed. Only commits in English were considered. Thus, our approach does not consider 
patches in any other language but English.
